Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2023-05-06 doc/manuscript/ESSDmanuscript.Rmd
check/edit through end of section 3.5

EFDB data submissions required information that was not recorded in previous versions of ForC, but rather was recorded in fields (`r table_ForCchanges`).
It was therefore necessary to return to original publications to retrieve relevant information, including 
(1) whether records of interest were presented in tables / text or digitized from figures (EFDB will not accept digitized data, whereas ForC does), 
(2) whether records of interest were presented directly, as opposed to having been calculated from related variables (for example, if a study presents aboveground biomass and root biomass but not total biomass, ForC would accept the sum of these as a valid estimate of total biomass, whereas EFDB would not)
(3) estimates in original units, 
and (4) confidence intervals (when not already in ForC).
We also checked that existing ForC records were complete and correct.

Manual review of records was the limiting step for data transfer to EFDB. 
We prioritized review of
(1) records from the Forest Global Earth Observatory [ForestGEO, @anderson-teixeira_ctfsforestgeo_2015; @davies_forestgeo_2021],
(2) studies with confidence intervals recorded in ForC (because uncertainty estimates are important to the IPCC),
(3) original publications containing large numbers of EFDB-relevant records,
and (4) records from tropical regions.
The latter criteria was motivated by the fact that although tropical forest is the single most important biome for climate change mitigation [@refs], ground-based data on tropical forest C cycling tend to be more scarce due to a variety of challenges [@delima_making_2022], and *tropical countries are more likely to apply Tier 1 methodology that bases forest C budgets on previously existing data [@ref].* <!-- I'm almost certain this is true, but need to confirm/ find ref.-->

@@

In addition to reviewing existing records, we added a total of `r n_records_added` new records to ForC.
These included **##** records from **##** studies [@piponiot_distribution_2022; @lutz_largediameter_2021; @refs<!--others?-->] that were not previously included in ForC.
In addition, we created new records for **##** EFDB-relevant estimates presented in the original publication that were not yet present in ForC.

>>>>

EFDB data submissions required information that was not recorded in previous versions of ForC, but for which new fields were created for EFDB compatibility (`r table_ForCchanges`).
It was therefore necessary to return to original publications to retrieve relevant information, including 
(1) estimates in original units, 
(2) confidence intervals (when not already in ForC),
(3) whether records of interest were presented in tables or text or digitized from figures (EFDB will not accept digitized data), 
(4) whether records of interest were presented directly, as opposed to having been calculated from related variables (for example, if a study presents aboveground biomass and root biomass but not total biomass, EFDB would not accept the sum of these as a valid record of total biomass)
We also checked that existing ForC records were complete and correct.

Manual review of records was the limiting step for data transfer to EFDB. 
We prioritized review of
(1) records from the Forest Global Earth Observatory [ForestGEO, @anderson-teixeira_ctfsforestgeo_2015; @davies_forestgeo_2021],
(2) studies with confidence intervals recorded in ForC (because uncertainty estimates are important to the IPCC),
(3) original publications containing large numbers of EFDB-relevant records,
and (4) records from tropical regions.
The latter criteria was motivated by the fact that although tropical forest is the single most important biome for climate change mitigation [@refs], ground-based data on tropical forest C cycling tends to be more scarce due to a variety of challenges [@refs; @delima_making_2022], and *tropical countries are more likely to apply Tier 1 methodology that bases forest C budgets on previously existing data [@ref].* <!-- I'm almost certain this is true, but need to confirm/ find ref.-->

@@

In addition to reviewing existing records, we added a total of `r n_records_added` new records to ForC.
These included 104 records<!--100 from Piponiot, 4 from Lutz--> from two studies [@piponiot_distribution_2022; @lutz_largediameter_2021] that were not previously included in ForC.
In addition, we created new records for `r n_records_added-104` EFDB-relevant estimates presented in the original publication that were not yet present in ForC.

### 2023-05-06 doc/manuscript/ESSDmanuscript.tex
check/edit through end of section 3.5

EFDB data submissions required information that was not recorded in
previous versions of ForC, but rather was recorded in fields (Appendix
A). It was therefore necessary to return to original publications to
retrieve relevant information, including (1) whether records of interest
were presented in tables / text or digitized from figures (EFDB will not
accept digitized data, whereas ForC does), (2) whether records of
interest were presented directly, as opposed to having been calculated
from related variables (for example, if a study presents aboveground
biomass and root biomass but not total biomass, ForC would accept the
sum of these as a valid estimate of total biomass, whereas EFDB would
not) (3) estimates in original units, and (4) confidence intervals (when
not already in ForC). We also checked that existing ForC records were
complete and correct.

Manual review of records was the limiting step for data transfer to
EFDB. We prioritized review of (1) records from the Forest Global Earth
Observatory
\citep[ForestGEO,][]{anderson-teixeira_ctfsforestgeo_2015, davies_forestgeo_2021},
(2) studies with confidence intervals recorded in ForC (because
uncertainty estimates are important to the IPCC), (3) original
publications containing large numbers of EFDB-relevant records, and (4)
records from tropical regions. The latter criteria was motivated by the
fact that although tropical forest is the single most important biome
for climate change mitigation \citep{refs}, ground-based data on
tropical forest C cycling tend to be more scarce due to a variety of
challenges \citep{delima_making_2022}, and \emph{tropical countries are
more likely to apply Tier 1 methodology that bases forest C budgets on
previously existing data \citep{ref}.}

@@

In addition to reviewing existing records, we added a total of 329 new
records to ForC. These included \textbf{\#\#} records from \textbf{\#\#}
studies
\citep{piponiot_distribution_2022, lutz_largediameter_2021, refs} that
were not previously included in ForC. In addition, we created new
records for \textbf{\#\#} EFDB-relevant estimates presented in the
original publication that were not yet present in ForC.

>>>>

EFDB data submissions required information that was not recorded in
previous versions of ForC, but for which new fields were created for
EFDB compatibility (Appendix A). It was therefore necessary to return to
original publications to retrieve relevant information, including (1)
estimates in original units, (2) confidence intervals (when not already
in ForC), (3) whether records of interest were presented in tables or
text or digitized from figures (EFDB will not accept digitized data),
(4) whether records of interest were presented directly, as opposed to
having been calculated from related variables (for example, if a study
presents aboveground biomass and root biomass but not total biomass,
EFDB would not accept the sum of these as a valid record of total
biomass) We also checked that existing ForC records were complete and
correct.

Manual review of records was the limiting step for data transfer to
EFDB. We prioritized review of (1) records from the Forest Global Earth
Observatory
\citep[ForestGEO,][]{anderson-teixeira_ctfsforestgeo_2015, davies_forestgeo_2021},
(2) studies with confidence intervals recorded in ForC (because
uncertainty estimates are important to the IPCC), (3) original
publications containing large numbers of EFDB-relevant records, and (4)
records from tropical regions. The latter criteria was motivated by the
fact that although tropical forest is the single most important biome
for climate change mitigation \citep{refs}, ground-based data on
tropical forest C cycling tends to be more scarce due to a variety of
challenges \citep{refs, delima_making_2022}, and \emph{tropical
countries are more likely to apply Tier 1 methodology that bases forest
C budgets on previously existing data \citep{ref}.}

@@

In addition to reviewing existing records, we added a total of 329 new
records to ForC. These included 104 records from two studies
\citep{piponiot_distribution_2022, lutz_largediameter_2021} that were
not previously included in ForC. In addition, we created new records for
225 EFDB-relevant estimates presented in the original publication that
were not yet present in ForC.
